---
title: "p8015_hw2_ax2173"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Problem 1

```{r read and clean the data}
subway_df = 
  read.csv("./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(route8 = as.character(route8), route9 = as.character(route9), route10 = as.character(route10), route11 = as.character(route11)) %>% 
  select(line, station_name, station_latitude, station_longitude, starts_with('route'), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, 'NO' = FALSE))
```

* There are some descriptions regarding these data:
  * The dataset contains these variables: `r colnames(subway_df)`
  * So far, I input the data, select some columns from the original dataset and convert the entry variable from character to a logical variable.
  * This dataset is a `r nrow(subway_df)` x `r ncol(subway_df)` dataset.

```{r calculate how many distinct stations}
station = 
  subway_df %>% 
  select(station_name, line) %>% 
  distinct
```

According the above code truck, we can get that there are `r nrow(station)` distinct stations.

```{r}
ada_count = 
  subway_df %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

According the above code truck, we can get that there are `r nrow(ada_count)` stations are ADA compliant.

```{r}
enty_without_vending = 
  subway_df %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

According the above code truck, we can get that there are `r enty_without_vending` stations are ADA compliant.

```{r}
serve_a = 
  subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = 'route',
    values_to = 'route'
  ) %>% 
  filter(route == 'A') %>% 
  select(station_name, line) %>% 
  distinct

serve_a_ada = 
    subway_df %>% 
  pivot_longer(
    route1:route11,
    names_to = 'route_number',
    names_prefix = 'route',
    values_to = 'route'
  ) %>% 
  filter(route == 'A', ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

According the above code truck, we can get that there are `r nrow(serve_a)` distinct stations serve the A train, and of the stations that serve the A train, `r nrow(serve_a_ada)` are ADA compliant.

## Problem 2

```{r read and clean the Mr. Trash Wheel sheet for P2}
mr_trash_df = 
  read_excel("./Data/Trash Wheel Collection Data.xlsx", skip = 1) %>% 
  janitor::clean_names() %>% 
  select(dumpster:homes_powered) %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         dumpster = as.double(dumpster),
         name = 'mr_trash_wheel')
```

```{r read and clean the Professor Trash Wheel for P2}
pro_trash_df = 
  read_excel("./Data/Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
         year = as.character(year),
         name = "professor_trash_wheel")
```

```{r combine these two dataset for P2}
trash_df = bind_rows(mr_trash_df, pro_trash_df)
```

* There are some descriptions regarding these data:
  * There are these varibales: `r colnames(trash_df)`
  * There are totally `r nrow(trash_df)` observations, and `r ncol(trash_df)` variables
  * The total weight of trash collected by Professor Trash Wheel is `r sum(pro_trash_df$weight_tons)` tons.
  * The total number of sport balls collected by Mr. Trash Wheel in 2020 is `r sum(mr_trash_df$sports_balls, mr_trash_df$year == 2020)` balls.

## Problem 3

```{r import data for P3}
pols_month_df = read_csv("./Data/fivethirtyeight_datasets/pols-month.csv")
unemployment_df = read_csv("./Data/fivethirtyeight_datasets/unemployment.csv")
snp_df = read_csv("./Data/fivethirtyeight_datasets/snp.csv")
```

```{r data clean for pols_month_df for P3}
pols_month_df = 
  pols_month_df %>% 
  separate(mon, into = c('year','month','day'), sep = '-') %>% 
  mutate(month = recode(month, '01' = 'Jan', '02' = 'Feb', '03' = 'Mar', '04' = 'Apr', '05' = 'May', '06' = 'Jun', '07' = 'Jul', '08' = 'Aug', '09' = 'Sep', '10' = 'Oct', '11' = 'Nov', '12' = 'Dec'),
         president = ifelse(pols_month_df$prez_gop == 0, 'gop', 'dem')) %>% 
  select(-day, -prez_dem, -prez_gop)
```

```{r data clean for snp_df for P3}
snp_df = snp_df %>% 
  separate(date, into = c('month','day','year'), sep = '/') %>% 
  mutate(month = recode(month, '1' = '01', '2' = '02', '3' = '03', '4' = '04', '5' = '05', '6' = '06', '7' = '07', '8' = '08', '9' = '09')) %>%
  arrange(year, month) %>% 
  mutate(month = recode(month, '01' = 'Jan', '02' = 'Feb', '03' = 'Mar', '04' = 'Apr', '05' = 'May', '06' = 'Jun', '07' = 'Jul', '08' = 'Aug', '09' = 'Sep', '10' = 'Oct', '11' = 'Nov', '12' = 'Dec')) %>%
  select(year, month, close)
```

```{r data clean for unemployment_df for P3}
unemployment_df = 
  unemployment_df %>% 
  pivot_longer(
    Jan : Dec,
    names_to = 'month',
    values_to = 'unemployment'
  ) %>% 
  janitor::clean_names() %>% 
  mutate(year = as.character(year))
```

```{r merge these dataframes for P3}
first_merge_df = left_join(pols_month_df, snp_df, by = c('year', 'month'))
resulting_df = left_join(first_merge_df, unemployment_df, by = c('year', 'month'))
```

* There are some descriptions regarding these data:
  * The pols_month dataset contains these variables: `r colnames(pols_month_df)`
  * The snp dataset contains these variables: `r colnames(snp_df)`
  * The unployment dataset contains these variables: `r colnames(unemployment_df)`
  * The resulting dataset is `r nrow(resulting_df)` x `r ncol(resulting_df)`, the range of years is `r range(resulting_df$year)`, and the final resulting dataset contains these variables: `r colnames(resulting_df)`


